---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {mldesign} - an R package for meaningful data splitting in applied machine learning

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
`r badger::badge_devel(color="blue")`
`r badger::badge_cran_release(color="orange")`
[![R-CMD-check](https://github.com/maxwestphal/mldesign/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/maxwestphal/mldesign/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/maxwestphal/mldesign/branch/main/graph/badge.svg)](https://app.codecov.io/gh/maxwestphal/mldesign?branch=main)
<!-- badges: end -->



The {mldesign} package allows meaningful data splitting for supervised machine learning tasks. 
The data splitting in {mldesign} is based on an **estimand** definition which is in turn based
on user-specified constraints.

**Constraints** can be thought of as generalized inclusion/exclusion criteria for 

1. Test observations
2. The relation between observations in train and test set
3. The training dataset(s)

In contrast to traditional techniques (hold-out, cross-validation, bootstrap), this leads to a
deterministic data splitting.

## Installation

You can install the development version of mldesign from [GitHub](https://github.com/) with:

```{r install, eval=FALSE}
# install.packages("devtools")
devtools::install_github("maxwestphal/mldesign")
```

## Getting started

An overview over important functions and classes and a minimal example can be found in the "overview" vignette:

```{r vignette, eval=FALSE}
library(mldesign)
vignette("overview", "mldesign")
```


## Roadmap

Planned features include:

1. Vast extension of traditional (random) data splitting methods 
2. Nested data splits
3. Export of data splitting to popular ML packages
3. print/summary/visualize functions for derived splits



