---
title: "overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mldesign)
```





## Main functionality

The {mldesign} package allows meaningful data splitting for supervised machine learning tasks. 
The data splitting in {mldesign} is based on an **estimand** definition which is in turn based
on user-specified constraints.

**Constraints** can be thought of generalized inclusion/exclusion criteria for 

1. Test observations
2. The relation between observations in train and test set
3. The training dataset(s)

In contrast to traditional techniques (hold-out, cross-validation, bootstrap), this leads to a
deterministic data splitting. The most important functions and classes are illustrated below:

```{r, setup2, include=FALSE}
library(nomnoml)
```

```{nomnoml, echo=FALSE, svg=TRUE, width="8in", height="8in"}
#direction: down
#stroke: black
#.frame: bold
#.obj: stroke=blue
#.fun: stroke=#ff8c00 visual=roundrect bold


[<frame> mldesign workflow for estimand-based data splitting|
  [<fun> constrain : function]
  [<fun> constrain : function] out -> [c1: mldesign_constraint]
  [<fun> constrain : function] out -> [c2: mldesign_constraint]
  [<fun> constrain : function] out -> [c3: mldesign_constraint]
  [c1: mldesign_constraint | name = "ic_age" | expr = "test$age >= 18" | vars = "age" | type = "population" | target = "test"]
  [c2: mldesign_constraint | name = "temporal_transferability" | expr = "test$year - train$year >= 1" | vars = "year" | type = "context" | target = "relation"]
  [c3: mldesign_constraint | name = "sufficient n_train" | expr = "nrow(train) >= 100" | vars = "" | type = "admissibility" | target = "train"]
  [c1: mldesign_constraint] in --> [<fun> specify_estimand : function]
  [c2: mldesign_constraint] in --> [specify_estimand : function]
  [c3: mldesign_constraint] in --> [specify_estimand : function] 
 
  [specify_estimand : function] out -> [my_estimand: mldesign_estimand | test = list(c1, ...) |  relation = list(c2, ...) |  train = list(c3, ...) ]
 
  [my_data: data.frame] in --> [<fun> derive_splits : function]
  [my_estimand: mldesign_estimand] in --> [derive_splits : function]
 
  [derive_splits : function] out -> [my_splits: mldesign_splits]
 
  [my_data: data.frame] in --> [<fun> get_data_split : function]
  [my_splits: mldesign_splits | list(\n list(train, test),\n list(train, test),\n ...\n) ] in --> [<fun> get_data_split : function]
  [index: integer] in --> [<fun> get_data_split : function]
  
  [get_data_split : function] out -> [my_data_split: list | train: data.frame | test: data.frame]
 
]
```

## Custom classes

1. **mldesign_constraint**: a single constraint
2. **mldesign_constraints**: a collection of constraints
3. **mldesign_estimand** (subclass of mldesign_spec): three constraints objects *test*, *relation*, *train*
4. **mldesign_splits**: a list of all train/test indices


## Important functions

1. **constrain()**: to define a constraint
2. **parse_constraint()**: to define a constraint only based on a single string
3. **combine_constraints()**: to combine multiple constraints
4. **specify_estimand()**: to define an estimand
5. **derive_splits(spec, data)**: to conduct the data splitting based on the estimand specification, returns indices only
6. **split_data(splits, data, split_id)**: to get a single train/set split, return data sets (subsets of original data)

## Overview

The following diagram shows important relationships between objects (of relevant classes) and functions
implemented in the {mldesign} package.





## Example

### Constraints

Generate some data for demonstration purposes
```{r}
set.seed(1337)
data <- generate_data(n_obs=2000)
names(data)
head(data)
```

Imagine we would want to apply our (trained) prediction model to predict the outcome variable only 
to patients older than 50 years:

```{r}
constrain(~ test$age > 50)
```
Note that variables in this expression will automatically be recognized as long as the syntax 
`test[['age']]` (preferred) or `test$age` (alternative, can cause problems in some cases) is
used. Currently, detection of variables is still in an early stage via `detect_vars()` 
but can be overwritten via (e.g.) `constrain(..., vars = "age")`.

We can define further (optional) information. This will not affect results down the line but 
will improve documentation
```{r}
constrain(~ test$age > 50, name="ic_age", type="population")
```

Note that we can specify the same constraint in a single character object with help of the parse_constraint function:

```{r}
"test[['age']] > 50 #name=ic_age #type=population" %>% parse_constraints()
```
Note that the use of the "#" character to specify further argument of constrain().

The function combine_constraints() (shortcut: cc()) can be used to bundle multiple mldesign_constraint
objects into a single mldesign_constraints object. 
It can also handle character inputs (via parse_constraint) or mldesign_constraints inputs.


### Estimand specification

To define a first, simple estimand, let's focus on the *relation* constraints. These constraints 
define the relationship between test and training observations. In this case, we specify constraints
that we expect to hold between training dataset **train** and testing dataset **test**. Note, that these
datasets do not yet exist explicitly but rather will be derived later 
(possibly in multiple versions) based on our estimand
specification and our entire dataset **data**.


```{r}
constraints_relation <- cc(
  constrain(~ (test[['year']] - train[['year']]) %in% 1:2,
            name="time_transferability", type="context"),
  constrain(~ test[['country']] != train[['country']],
            name="region_transferability", type="context")
) 
```

```{r}
constraints_relation %>% str(2)
```
```{r}
my_estimand <- specify_estimand(constraints = constraints_relation,
                                name = "my_estimand")
my_estimand
```




### Data splitting 

Now, we use the estimand to conduct (deterministically) the corresponding data splitting:

```{r}
splits <- derive_splits(my_estimand, data) 
```

```{r}
length(splits)
str(splits[[1]])
```
As we can see, we obtain a list of train/test indices corresponding to the estimand specification.
To finally obtain concrete train/test sets (the first one out of the overall list):

```{r}
split_data(data, splits, idx=1) %>% str()
```



